Exploring the composition of lithic assemblages in Mesolithic south-eastern Norway

### 2021-04-25 analysis/paper/paper.Rmd
Exploring the data some more - setting up for correlation plots.

```{r libraries, eval = TRUE}
library(tidyverse)
library(patchwork)
library(gt)
library(FactoMineR)
library(factoextra)
library(sf)
library(rnaturalearth)
```

@@

# Repeat but for flint only
flint2 <- sites %>% 
  select(-n, -beach_flint_flint_nodule, -hammerstone, -excavation_project, 
         -reported_start_bp, -reported_end_bp, -askeladden_id) %>% 
  select(!ends_with("other")) %>% 
  mutate(chip_freq =  sum(chip) / artefact_count,
         core_freq = sum(core) / artefact_count,
         blanks_freq = sum( blade, microblade, fragment, flake) / artefact_count,
         secondary_freq =  sum(axe_debitage, axe, chisel, 
                microlith, tanged_point, transverse_point, single_edged_point,
                hognipen_point, projectile_undefined, microburin, scraper, 
                knife, burin, drill, blade_retouch, microblade_retouch, 
                chip_retouch, fragment_retouch, flake_retouch) / artefact_count,
         flint_dens = flint_count / excavated_m3) %>%
  select(-blanks_flint, -secondary_flint, -excavated_m3, -artefact_count,
         -flint_count, -nonflint_count, -grinding_slab_stone, -core, 
         -core_fragment, -chip) %>% 
  column_to_rownames("site")

wabicor <- cor(log10(lithics2$lithic_dens), log10(lithics2$secondary_freq))
wabicorflint <- cor(log10(flint2$flint_dens), log10(flint2$secondary_freq))
```

@@

```{r pca-analysis, fig.cap = "PCA.", fig.height = 7, fig.width = 10}
pcalithics <- PCA(select(lithics2, -mean_date), scale.unit = TRUE, graph = FALSE)
fviz_pca(pcalithics, geom.ind = "point", fill.ind = lithics2$mean_date / 1000, 
          repel = TRUE, col.var = "black", pointshape = 21,  pointsize = 2.9,
         title = "") + 
  scale_fill_gradient2(name = "ka BP",  low="yellow", mid="orange", 
                        high="red", midpoint = centdate) +
  theme(legend.position = "bottom")

>>>>

```{r libraries, eval = TRUE}
library(tidyverse)
library(patchwork)
library(gt)
library(FactoMineR)
library(factoextra)
library(sf)
library(rnaturalearth)
```
```{r get-functions, include = FALSE}
#source("plot_functions.R", local = knitr::knit_global())
```

@@

lithics3 <- sites %>% 
  select(-n, -beach_flint_flint_nodule, -hammerstone, -excavation_project,
         -reported_start_bp, -reported_end_bp, -askeladden_id) %>% 
  # The reason for not using .keep below is for the option to 
  #retain artefact_count
  mutate(chip_freq =  sum(chip, chip_other),
         core_freq = sum(core, core_other),
         blanks = sum(blade, microblade, fragment, flake, flake_other, 
                     fragment_other, blade_other,  microblade_other, -core, -core_fragment, -chip, 
         -chip_other, -core_other, -core_fragment_other),
         secondary =  sum(axe_debitage, axe, chisel, 
                microlith, tanged_point, transverse_point, single_edged_point,
                hognipen_point, projectile_undefined, microburin, scraper, 
                knife, burin, drill, blade_retouch, microblade_retouch, 
                chip_retouch, fragment_retouch, flake_retouch, axe_other,
                axe_debitage_other, chisel_other, club_hatchet_other, 
                grinding_slab_stone, projectile_point_other, knife_other,
                scraper_other, drill_other, microblade_retouch_other, 
                blade_retouch_other, fragment_retouch_other,
                flake_retouch_other),
         lithic_dens = artefact_count / excavated_m3) %>%
  select(-blanks_lithics, -secondary_lithics, -excavated_m3, -artefact_count,
         -flint_count, -nonflint_count, -core, -core_fragment, -chip, 
         -chip_other, -core_other, -core_fragment_other) %>% 
  column_to_rownames("site")

# Repeat but for flint only
flint2 <- sites %>% 
  select(-n, -beach_flint_flint_nodule, -hammerstone, -excavation_project, 
         -reported_start_bp, -reported_end_bp, -askeladden_id) %>% 
  select(!ends_with("other")) %>% 
  mutate(chip_freq =  sum(chip) / artefact_count,
         core_freq = sum(core) / artefact_count,
         blanks_freq = sum(blade, microblade, fragment, flake) / artefact_count,
         secondary_freq =  sum(axe_debitage, axe, chisel, 
                microlith, tanged_point, transverse_point, single_edged_point,
                hognipen_point, projectile_undefined, microburin, scraper, 
                knife, burin, drill, blade_retouch, microblade_retouch, 
                chip_retouch, fragment_retouch, flake_retouch) / artefact_count,
         flint_dens = flint_count / excavated_m3) %>%
  select(-blanks_flint, -secondary_flint, -excavated_m3, -artefact_count,
         -flint_count, -nonflint_count, -grinding_slab_stone, -core, 
         -core_fragment, -chip) %>% 
  column_to_rownames("site")

wabicor <- cor.test(log10(lithics2$lithic_dens), log10(lithics2$secondary_freq))
wabicorflint <- cor(log10(flint2$flint_dens), log10(flint2$secondary_freq))
```

@@

```{r pca-analysis, fig.cap = "PCA.", fig.height = 7, fig.width = 10}
pcalithics <- PCA(select(lithics2, -mean_date, -nonflint_freq), scale.unit = TRUE, graph = FALSE)
fviz_pca(pcalithics, geom.ind = "point", fill.ind = lithics2$mean_date / 1000, 
          repel = TRUE, col.var = "black", pointshape = 21,  pointsize = 2.9,
         title = "") + 
  scale_fill_gradient2(name = "ka BP",  low="yellow", mid="orange", 
                        high="red", midpoint = centdate) +
  theme(legend.position = "bottom")
